Clustering time series based on probability distributions across temporal granularities

### 2021-11-18 main.Rmd
validation proofread and positioning table for improved reading

The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities. Consider a group that has a different weekday and weekend behavior in the summer but not in the winter. This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (weekend/weekday and month-of-year in this example). As a result, in this scenario, we analyze a combination of categories generated from different distributions. Display of design and related results can be found in supplementary.

@@

Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns). This can be used to determine the number of clusters for each scenario. When $sindex$ for each scenario are examined, it appears that $k = 5, 4, 4$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$. Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.

@@

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter. The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.  Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure is related to granularities.

>>>>

The proposed methods could be extended when two granularities of interest interact and we want to group subjects based on the interaction of the two granularities. Consider a group that has a different weekday and weekend behavior in the summer but not in the winter. This type of combined behavior across granularities can be discovered by evaluating the distribution across combinations of categories for different interacting granularities (weekend/weekday and month-of-year in this example). As a result, in this scenario, we analyze a combination of categories generated from different distributions. Display of design and related results can be found in supplementary material.

@@

Figure \ref{fig:sindex-plot-validation} shows $sindex$ plotted against the number of clusters ($k$) for the range of mean differences (rows) under the different scenarios (columns). This can be used to determine the number of clusters for each scenario. When $sindex$ for each scenario are examined, it appears that $k = \{5, 4, 4\}$ is justified for scenarios S1, S2, and S3, respectively, given the sharp decrease in $sindex$ from that value of $k$. Thus, the number of clusters corresponds to the number of designs that were originally considered in each scenario.

@@

Clustering with the new distances is illustrated on the smart meter energy usage for a sample of customers from @smart-meter. The full data contains half-hourly general supply in KwH for 13,735 customers, resulting in 344,518,791 observations in total.<!--In most cases, electricity data is expected to have multiple seasonal patterns like daily, weekly or annual. We do not learn about these repetitive behaviors from the linear view because too many measurements all squeezed in that representation. Hence we transition into looking at cyclic granularities, that can potentially provide more insight on their repetitive behavior.--> The raw data for these consumers is of unequal length, with varying starting and end dates.  Additionally, there were missing values in many series. (The supplementary material contains details from checking for systematic missingness.) Because our proposed methods evaluate probability distributions rather than raw data, these data issues are not problematic, unless there is any systematic structure related to granularities.
